---
title: "RepData_PeerAssessment1"
author: "Wang Weiqiang"
date: "Friday, March 14, 2015"
output: html_document
---


## Loading and preprocessing the data

```{r}
srouceData<-read.csv("activity.csv")
tidyData<-na.omit(srouceData)
```


## What is mean total number of steps taken per day?
```{r}
stepsPerDay=aggregate(tidyData$steps, by=list(substr(tidyData$date,1,10)), FUN=sum)
colnames(stepsPerDay)<-c("Day","Steps")
stepsPerDay
```

```{r}
barplot(stepsPerDay$Steps,axisnames=TRUE,names.arg=stepsPerDay$Day,main="Total number of steps taken per day",xlab="Date",ylab="Numbers of Steps")
```

```{r}
mean(stepsPerDay$Steps)
median(stepsPerDay$Steps)
```

## What is the average daily activity pattern?
```{r}
numberOfDays=length(summary(tidyData$date))
activityPatternData=aggregate(tidyData$steps/numberOfDays, by=list(tidyData$interval), FUN=mean)
colnames(activityPatternData)<-c("interval","Steps")
plot(activityPatternData,type="l")
```

###Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxNumberOfSteps=max( activityPatternData$Steps)
activityPatternData[activityPatternData$Steps ==maxNumberOfSteps,]
```
## Imputing missing values
### total number of missing values in the dataset
```{r}
targetColumn <- srouceData$steps
nrow(srouceData[is.na(targetColumn),])
```
### filling in all of the missing values
```{r}
newData<-srouceData
stepsPer5minute=as.integer(sum(tidyData$steps)/nrow(tidyData))
for (i in 1:nrow(newData)){
  if(is.na(newData[i,]$steps)){
    newData[i,]$steps=stepsPer5minute
  }
}
```

### histogram of the total number of steps taken each day

```{r}
stepsPerDay=aggregate(newData$steps, by=list(substr(newData$date,1,10)), FUN=sum)
colnames(stepsPerDay)<-c("Day","Steps")
barplot(stepsPerDay$Steps,axisnames=TRUE,names.arg=stepsPerDay$Day,main="Total number of steps taken per day",xlab="Date",ylab="Numbers of Steps")
```

### mean total number of steps taken per day

```{r}
mean(stepsPerDay$Steps)
```

### median total number of steps taken per day

```{r}
median(stepsPerDay$Steps)
```
These values are differ from the estimates from the first part of the assignment.
the mean and mdian values are lower than part 1, because the sampling frequency are higher.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
newData["DayType"]<-NA
for (i in 1:nrow(newData)){
  
    date <- as.POSIXlt(newData[i,]$date,format="%Y-%m-%d")
    if(date$wday==0 | date$wday==6){
      newData[i,]$DayType<-"weekend"
    }
    else
    {
      newData[i,]$DayType<-"weekday"
    }
  
}

weekendData=subset(newData,DayType=="weekend")
weekdaysData=subset(newData,DayType=="weekday")


weekendActivity=aggregate(weekendData$steps/length(summary(weekendData$date)), by=list(weekendData$interval), FUN=mean)
colnames(weekendActivity)<-c("interval","Steps")

weekdaysActivity=aggregate(weekdaysData$steps/length(summary(weekdaysData$date)), by=list(weekdaysData$interval), FUN=mean)
colnames(weekdaysActivity)<-c("interval","Steps")

par(mfrow=c(2,1))
plot(weekendActivity,type="l",xlab="weekend",ylab="Number of steps")
plot(weekdaysActivity,type="l",xlab="weekday",ylab="Number of steps")


```
